Does It Scale? Static Output Approximation of PHP Web Applications Using Symbolic Execution

### 2017-01-11 paper.tex
added patrs of 3

\setlength\parindent{0pt}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{blindtext}
\usepackage{booktabs}
\usepackage{paralist}
\usepackage{listings,multicol}
\usepackage{lipsum}
%\usepackage{cite}
\usepackage{enumerate}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{ucs}
\usepackage{framed}

@@

\lstset{
	language=PHP,
	extendedchars = \true,
	basicstyle=\ttfamily\footnotesize,
	basewidth={0.5em,0.4em}, % for other columns settings
	commentstyle = \color{gray},
	columns=fullflexible,
	numbers=left,
	numberstyle=\tiny,
	frame=tb,
	showstringspaces=false,
	xleftmargin=2em,
	framexleftmargin=1.5em,
	backgroundcolor=\color{lightgrey},
	inputencoding = utf8x,
    keepspaces = true,
    keywordstyle = \bfseries,
}

\newcommand{\dollar}{\mbox{\textdollar}}

@@

Based on the approximated output representation subsequent analyses can be
conducted, which have been implemented in the following tools:

\begin{itemize}
	\item {\emph{PhpSync}: Using all variants of  HTML client page output, every
	single one can be statically checked for Markup Validity, i.e., if the web page
	conforms to syntactical specifications for HTML and other client-side
	languages. After validation errors are detected, auto-fixes are provided.
	Otherwise, the validation error can be traced back to source code responsible
	for the defect, which then can be refactored manually
	\cite{Nguyen:2011:AFH:2190078.2190142}. }

@@

\subsection{Towards Scalable Output Approximation?}
As our goal is to have practical tools support, we approach this question by
re-implementing the symbolic execution semantics described in
\cite{Nguyen:2014:BCG:2635868.2635928} and extending it with support for
additional language features. We choose to aim to support symbolic execution
for a real-world example system of reasonable size that incorporates the
missing language features described in Section \ref{sec:limitations} and is
still developed and maintained: \textsf{WordPress} is a popular open-source
Content Management System (CMS) with around 300k lines that provides a vivid
plug-in environment and is maintained by a large community.
For our implementation of a symbolic execution engine for PHP, \textsf{Oak}, we
have chosen a test-driven approach with continuous integration using regression
tests based on \textsf{WordPress} and \textsf{SchoolMate} (see Section
\ref{sec:experiment_setup}). Oak incorporates the existing execution semantics
of \cite{Nguyen:2014:BCG:2635868.2635928} and extends it with support for
object-oriented orientation (class definition, class instantiation and method
invocation) as well as PHP array operations. A list of the functions provided
by the PHP standard library can be found at \cite{PHP}.
Throughout the implementation we did not achieve high code coverage for
WordPress, which was the main motivation to further investigate limitations in
the wild.

\subsection{Things We Learned The Hard Way}
\textcolor{blue}{\blindtext}

\subsubsection{Symbolic vs Concrete Execution}
\textcolor{blue}{\blindtext}

@@

\subsection{Experiment Setup} \label{sec:experiment_setup}
For the evaluation we symbolically executed a selection of PHP systems. We
selected a corpus of twelve PHP systems with regard to system size and recency
as the corpus of the case study for \cite{Nguyen:2014:BCG:2635868.2635928} only
contained small-scale systems that are not maintained any more. The full list
of PHP systems is shown in Table \ref{corpus}. Our selection of PHP systems
includes

%
\begin{itemize}
	\item four small-scale systems that selected from the previously mentioned case
	study in order to compare our results with the previous symbolic execution
	engine \cite{Nguyen:2014:BCG:2635868.2635928},

@@

\section{Related Work}
\textcolor{blue}{\blindtext\blindtext}

%References
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

>>>>

% just in case
\usepackage{etex}

% does not seem to work with sigconf template
%\usepackage{cite}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{blindtext}
\usepackage{booktabs}
\usepackage{paralist}
\usepackage{listings,multicol}
\usepackage{lipsum}

\usepackage{enumerate}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{ucs}
\usepackage{framed}

\setlength\parindent{0pt}

@@

\lstset{
	language=PHP,
	extendedchars = \true,
	basicstyle=\ttfamily\footnotesize,
	basewidth={0.5em,0.4em}, % for other columns settings
	commentstyle = \color{gray},
	columns=fullflexible,
	numbers=left,
	numberstyle=\tiny,
	frame=tb,
	showstringspaces=false,
	xleftmargin=2em,
	framexleftmargin=1.5em,
	backgroundcolor=\color{lightgrey},
	inputencoding = utf8x,
    keepspaces = true,
    keywordstyle = \bfseries
}

%\newcommand{\dollar}{\mbox{\textdollar}}

@@

Based on the approximated output representation subsequent analyses can be
conducted, which have been implemented in the following tools:
\begin{itemize}
	\item {\emph{PhpSync}: Using all variants of  HTML client page output, every
	single one can be statically checked for Markup Validity, i.e., if the web page
	conforms to syntactical specifications for HTML and other client-side
	languages. After validation errors are detected, auto-fixes are provided.
	Otherwise, the validation error can be traced back to source code responsible
	for the defect, which then can be refactored manually
	\cite{Nguyen:2011:AFH:2190078.2190142}. }

@@

\subsection{Towards Scalable Output Approximation?}
As our goal is to have practical tools support, we approach this question by
re-implementing the symbolic execution semantics described in
\cite{Nguyen:2014:BCG:2635868.2635928} and extending it with support for
additional language features. We choose to aim to support symbolic execution
for a real-world example system of reasonable size that incorporates the
missing language features described in Section \ref{sec:limitations} and is
still developed and maintained: \textsf{WordPress} is a popular open-source
Content Management System (CMS) with around 300k lines that provides a vivid
plug-in environment and is maintained by a large community.
For our implementation of a symbolic execution engine for PHP,
\textsf{Oak}\footnote{The implementation can be found
at \url{github.com/smba/oak}.}, we have chosen a test-driven approach with
continuous integration using regression tests based on \textsf{WordPress} and \textsf{SchoolMate} (see Section \ref{sec:experiment_setup}). Oak incorporates the existing execution semantics
of \cite{Nguyen:2014:BCG:2635868.2635928} and extends it with support for
object-oriented orientation (class definition, class instantiation and method
invocation) as well as PHP array operations. A list of the functions provided
by the PHP standard library can be found at \cite{PHP}.
Throughout the implementation we did not achieve high code coverage for
WordPress, which was the main motivation to further investigate limitations in
the wild.

\subsection{Experience Report}
This sections sums up 
\textcolor{blue}{\blindtext}

\subsubsection{Symbolic vs Concrete Execution}
Throughout the implementation, we basically followed the symbolic execution
semantics described in \cite{Nguyen:2014:BCG:2635868.2635928}. Nevertheless, for additional language features as
well as standard library functions we experienced it to be beneficial to the
approximation to execute them as concretely as possible.
For instance, \sf{WordPress} uses a text filter mechanism, where different
filter functions are applied sequentially to a string literal. All filters can
be linked to a hook using  associative arrays so that when the hook is
triggered, all filters linked to it are applied. As described in Section \ref{}
we execute loops, in particular while loops, only once due to an possibly
indeterminable number of iterations. PHP yet offers more loop constructs: the
do-while loop and the foreach loop. While for the do-while loop as well as for
the while loop the number of iterations is possibly indeterminable, the foreach
loop iterates over all items of a given array whose number of items is known at
runtime. For \sf{WordPress}, we are able to achieve a more accurate
approximation and higher code coverage results by concretely executing foreach
loops as more server-side code (filter functions in our case) are accessed and
executed. In turn, exhaustively executing foreach loops and filter functions
concretely is a trade-off between accuracy and performance, as single hook can
be called for different contexts as we are exploring multiple execution paths.

An additional aspect on the boundary of concrete and symbolic execution, or
effort and accuracy respectively, is the extent to which program inputs are
assumed to be symbolic. As shown in Section \ref{} introducing symbolic values
can be problematic if they are propagated to language features requiring
concrete information, such as include expressions or indirect function calls.
We manually scaled the number of sources of symbolic inputs by providing static
dummy inputs such as a concrete form input, file name or configuration entry.
For specific cases, where concrete information is inevitably required, we can
achieve higher code coverage and a more accurate approximation, as more code
becomes accessible. Nevertheless, while detecting dynamic features can be
automated, it requires manual inspection and understanding of the web
application to provide appropriate dummy input data. Again, this is a trade-off
between the effort to provide input data and accuracy of the approximation and
code coverage respectively.

@@

\subsection{Experiment Setup} \label{sec:experiment_setup}
For the evaluation we symbolically executed a selection of PHP systems. We
selected a corpus of twelve PHP systems with regard to system size and recency
as the corpus of the case study for \cite{Nguyen:2014:BCG:2635868.2635928} only
contained small-scale systems that are not maintained any more. The full list
of PHP systems is shown in Table \ref{corpus}. Our selection of PHP systems
includes%
\begin{itemize}
	\item four small-scale systems that selected from the previously mentioned case
	study in order to compare our results with the previous symbolic execution
	engine \cite{Nguyen:2014:BCG:2635868.2635928},

@@

\section{Related Work}
\cite{Hills:2013:ESP:2483760.2483786}
\cite{wassermann2007sound,minamide2005static}
\textcolor{blue}{\blindtext\blindtext}

\newpage
%References
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}
